---
title: "T_c approach: <br> Affinity estimations"
author: "Vicente J. Ontiveros"
date: "5/19/2022"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = F, message = F)
```

```{r, warning = F}

library(tidyverse)

# Loading subsampling 10
load("analysis/vicente/ss10.RData")

ss10.est <- data.table::rbindlist(parall)
ss10.est <- ss10.est %>% add_column(Subsampling = 10)

# Loading subsampling 25
load("analysis/vicente/ss25.RData")

ss25.est <- data.table::rbindlist(parall)
ss25.est <- ss25.est %>% add_column(Subsampling = 25)

# Loading subsampling 50
load("analysis/vicente/ss50.RData")

ss50.est <- data.table::rbindlist(parall)
ss50.est <- ss50.est %>% add_column(Subsampling = 50)

# Loading subsampling 75
load("analysis/vicente/ss75.RData")

ss75.est <- data.table::rbindlist(parall)
ss75.est <- ss75.est %>% add_column(Subsampling = 75)

# Loading complete sampling
load("analysis/vicente/estimates_neutral_v2.RData")

neutral <- neutral %>% add_column(Subsampling = 100)

# Load metadata
n_metadata <- read.csv("data/simulations/neutral_metadata_v2.csv")

# Gathering all estimates -------------------------------------------------
neutral.est <- rbind(neutral, ss75.est, ss50.est, ss25.est, ss10.est)

short.metadata <- n_metadata %>% select(2, 6, 7, 8)

neutral.res <- inner_join(neutral.est, short.metadata)

save(neutral.res, file = "analysis/vicente/Affinity_subsamplings.RData")
```

# Affinity

## What is affinity?

A measure of instantaneous change in community composition. Depends on:

- The asymptote, $V_{Max}$ in Michaelis-Menten dynamics, assymptotic dissimilarity in ETIB.
- Half-saturation, $K_M$ in M-M, a measure related to T_c in ETIB. 

## Same affinity...

```{r}
library(tidyverse)

col <- 0.1
ext <- 0.2

assymptotic <- 1 - col/(col + 2 * ext)
half <- log((2 * col + 3 * ext)/(col + 2 * ext))/(col + ext)

x <- 0:100/5

y <- assymptotic * x / (half + x)

y2 <- 1 - (col + ext * exp(-x * (col + ext)))/(col + 2 * ext - ext * exp(-x * (col + ext)))

df1 <- data.frame(Time = x, Dissimilarity = y, Type = "Michaelis-Menten")
df2 <- data.frame(Time = x, Dissimilarity = y2, Type = "ETIB")

df <- rbind(df1, df2)

ggplot(df, aes(x = Time, y = Dissimilarity, color = Type)) + 
  geom_line(size = 1) + 
  scale_color_manual(values = c("#EF3054", "#0b6e4f")) +
  theme_bw() +
  geom_hline(yintercept = assymptotic, linetype = 2) + 
  annotate(geom = "text", x = 4.56, y = .4, label = "Half - saturation time") +
  annotate(geom = "text", x = 2, y = .77, label = "Assymptotic Jaccard") + 
  theme(aspect.ratio = .618, panel.grid.minor = element_blank(), legend.position = "bottom")

```

... but different temporal dynamics!

## Sampling completeness

```{r}
ggplot(neutral.est, 
       aes(x = Subsampling, y = raw_Affinity, group = (Subsampling), color = as.factor(Subsampling))) +
   geom_boxplot() +
   theme_bw() +
   theme(legend.position = "none", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

Very slow change in the true community!

## Duration

```{r}
ggplot((neutral.est), aes(x = Duration, y = raw_Affinity, color = as.factor(Subsampling))) +
   geom_point(alpha = .7) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

Slight but significant tendency to decrease with duration. Asymptotically unbiased.

## Duration - 100%

```{r}
ggplot(neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000), aes(x = Duration, y = raw_Affinity)) +
   geom_point(alpha = .9, color = "#E76BF3") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000),
           raw_Affinity ~ Duration))
```

Almost significant tendency to decrease with duration. Asymptotically unbiased. $\Theta$ = 40, M = .2, N = 1000.

## Species pool size

```{r}
ggplot(neutral.res %>% filter(parameter_id < 10), aes(x = THETA, y = raw_Affinity, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

## Species pool size - 100% - duration 13

```{r}
ggplot(neutral.res %>% 
          filter(parameter_id < 10) %>% 
          filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = THETA, y = raw_Affinity, group = THETA)) +
    geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   geom_jitter(alpha = .25) +
  theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id < 10) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13),
           raw_Affinity ~ THETA))
```

Very small but significant increase. N = 1000, M = .2.

## Local community size

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20), aes(x = N, y = raw_Affinity, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

## Local comm. size - 100% - duration 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = N, y = raw_Affinity, group = N)) +
       geom_boxplot(outlier.shape = NA, color = "#E76BF3") +
  geom_jitter(alpha = .25) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

$\Theta$ = 40, M = .2.

## Rate of new species arrival

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% filter(parameter_id < 21), 
       aes(x = M, y = raw_Affinity, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

## New species arrival - 100% - dur. 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% 
          filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
       filter(Duration == 13), 
       aes(x = M, y = raw_Affinity, group = M)) +
   geom_jitter(alpha = .25) +   
   geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 9) %>% 
              filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
              filter(Duration == 13),
           raw_Affinity ~ M))
```

Very small but significant increase. $\Theta$ = 40, N = 1000.

# $T_c$

$T_c$ only measures the speed of change, not its extent.

## Sampling completeness

```{r}
ggplot(neutral.est, 
       aes(x = Subsampling, y = raw_T_c, group = (Subsampling), color = as.factor(Subsampling))) +
   geom_boxplot() +
   theme_bw() +
   theme(legend.position = "none", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

Very slow change in the true community!

## Duration

```{r}
ggplot((neutral.est), aes(x = Duration, y = raw_T_c, color = as.factor(Subsampling))) +
   geom_point(alpha = .7) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```



## Duration - 100%

```{r}
ggplot(neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000), aes(x = Duration, y = raw_T_c)) +
   geom_point(alpha = .9, color = "#E76BF3") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000),
           raw_T_c ~ Duration))
```

$\Theta$ = 40, M = .2, N = 1000.

## Summary
- Difficult to see things maybe because slow pace of change. 
- Increasing species pool size while maintaining local community size and migration constant reduces $T_c$ (less individuals per species increases the pace of change).
- Local community size seems to have no effect on $T_c$ (related to the abundance distribution?).
- Increasing the rate of new species arrival reduces $T_c$ (more migration increases the pace of change).
- See figures in More Slides: T_c.


# Characteristic Jaccard

## Sampling completeness

```{r}
ggplot(neutral.est, 
       aes(x = Subsampling, y = raw_J_t_c, group = (Subsampling), color = as.factor(Subsampling))) +
   geom_boxplot() +
   theme_bw() +
   theme(legend.position = "none", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

Small change in the true community!

## Duration

```{r}
ggplot((neutral.est), aes(x = Duration, y = raw_J_t_c, color = as.factor(Subsampling))) +
   geom_point(alpha = .7) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

Biased estimator under neutral, zero-sum dynamics? 

## Duration - 100%

```{r}
ggplot(neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000), aes(x = Duration, y = raw_J_t_c)) +
   geom_point(alpha = .9, color = "#E76BF3") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(Subsampling == 100) %>% filter(THETA == 40) %>% 
          filter(M == .2) %>% filter(N == 1000),
           raw_J_t_c ~ Duration))
```

$\Theta$ = 40, M = .2, N = 1000.

## Summary
- Unreliable estimator because of very limited change?
- Estimates seem to fail.
- Contradictory results for species pool size, migration.
- Biases cancel each other using affinity?
- See figures in More Slides: Characteric Jaccard.

# More Slides: Characteristic Jaccard.

## Species pool size

```{r}
ggplot(neutral.res %>% filter(parameter_id < 10), aes(x = THETA, y = raw_J_t_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

## Species pool size - 100% - duration 13

```{r}
ggplot(neutral.res %>% 
          filter(parameter_id < 10) %>% 
          filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = THETA, y = raw_J_t_c, group = THETA)) +
    geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   geom_jitter(alpha = .25) +
  theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") 
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id < 10) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13),
           raw_J_t_c ~ THETA))
```

Small but significant decrease! N = 1000, M = .2.

## Local community size

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20), aes(x = N, y = raw_J_t_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

## Local comm. size - 100% - duration 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = N, y = raw_J_t_c, group = N)) +
       geom_boxplot(outlier.shape = NA, color = "#E76BF3") +
  geom_jitter(alpha = .25) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```



```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 20) %>% 
              filter(Subsampling == 100) %>% 
              filter(Duration == 13),
           raw_J_t_c ~ N))
```

Almost significant decrease. $\Theta$ = 40, M = .2.

## Rate of new species arrival

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% filter(parameter_id < 21), 
       aes(x = M, y = raw_J_t_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

## New species arrival - 100% - dur. 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% 
          filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
       filter(Duration == 13), 
       aes(x = M, y = raw_J_t_c, group = M)) +
   geom_jitter(alpha = .25) +   
   geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic Jaccard") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 9) %>% 
              filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
              filter(Duration == 13),
           raw_J_t_c ~ M))
```

Significant decrease. $\Theta$ = 40, N = 1000.


# More Slides: T_c.

## Species pool size

```{r}
ggplot(neutral.res %>% filter(parameter_id < 10), aes(x = THETA, y = raw_T_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

## Species pool size - 100% - duration 13

```{r}
ggplot(neutral.res %>% 
          filter(parameter_id < 10) %>% 
          filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = THETA, y = raw_T_c, group = THETA)) +
    geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   geom_jitter(alpha = .25) +
  theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") 
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id < 10) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13),
           raw_T_c ~ THETA))
```

Small but significant decrease. N = 1000, M = .2.

## Local community size

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20), aes(x = N, y = raw_T_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

## Local comm. size - 100% - duration 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20) %>% filter(Subsampling == 100) %>% 
          filter(Duration == 13), 
       aes(x = N, y = raw_T_c, group = N)) +
       geom_boxplot(outlier.shape = NA, color = "#E76BF3") +
  geom_jitter(alpha = .25) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```



```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 20) %>% 
              filter(Subsampling == 100) %>% 
              filter(Duration == 13),
           raw_T_c ~ N))
```

No linear relationship. $\Theta$ = 40, M = .2.

## Rate of new species arrival

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% filter(parameter_id < 21), 
       aes(x = M, y = raw_T_c, color = as.factor(Subsampling))) +
   geom_jitter(alpha = .9) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

## New species arrival - 100% - dur. 13

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% 
          filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
       filter(Duration == 13), 
       aes(x = M, y = raw_T_c, group = M)) +
   geom_jitter(alpha = .25) +   
   geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Characteristic time") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 9) %>% 
              filter(parameter_id < 21) %>% filter(Subsampling == 100) %>% 
              filter(Duration == 13),
           raw_T_c ~ M))
```

Significant decrease. $\Theta$ = 40, N = 1000.




# Old slides
## Species pool size - complete

```{r}
ggplot(neutral.res %>% filter(parameter_id < 10) %>% filter(Subsampling == 100), 
       aes(x = THETA, y = raw_Affinity, group = THETA)) +
    geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   geom_jitter(alpha = .1) +
  theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id < 10) %>% filter(Subsampling == 100),
           raw_Affinity ~ THETA))
```

Very small but significant increase.

## Local community size - complete

```{r}
ggplot(neutral.res %>% filter(parameter_id > 20) %>% filter(Subsampling == 100), 
       aes(x = N, y = raw_Affinity, group = N)) +
       geom_boxplot(outlier.shape = NA, color = "#E76BF3") +
  geom_jitter(alpha = .1) +
   scale_x_log10() +
   annotation_logticks(sides = "b") +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```



## Rate of new species arrival - complete

```{r}
ggplot(neutral.res %>% filter(parameter_id > 9) %>% filter(parameter_id < 21) %>% filter(Subsampling == 100), 
       aes(x = M, y = raw_Affinity, group = M)) +
   geom_jitter(alpha = .1) +   
   geom_boxplot(outlier.shape = NA, color = "#E76BF3", fill = NA) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.minor = element_blank(), aspect.ratio = .618) + 
   ylab("Affinity") +
   labs(color = "Subsampling")
```

```{r, include = F}
summary(lm(data = neutral.res %>% filter(parameter_id > 9) %>% filter(parameter_id < 21) %>% filter(Subsampling == 100),
           raw_Affinity ~ M))
```

Very small but significant increase.


# End

<!-- ## R Markdown -->

<!-- This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. -->

<!-- When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. -->

<!-- ## Slide with Bullets -->

<!-- - Bullet 1 -->
<!-- - Bullet 2 -->
<!-- - Bullet 3 -->

<!-- ## Slide with R Output -->

<!-- ```{r cars, echo = TRUE} -->
<!-- summary(cars) -->
<!-- ``` -->

<!-- ## Slide with Plot -->

<!-- ```{r pressure} -->
<!-- plot(pressure) -->
<!-- ``` -->

