---
title: "Time-series analysis - T_c approach"
output:
  pdf_document: default
  html_notebook: default
editor_options:
  chunk_output_type: inline
---

## Take-home messages.

1. Compositional change is too fast in the simulations. 
2. Some combinations of parameters are too extreme (20 sp. 20000 individuals, or 200 species 2000 individuals).
3. If we do not take into account our previous knowledge of the species pool, colonization-extinction rates become biased at small durations and, so, the $T_c$ approach. However, the estimator is asymptotically unbiased and effective, meaning that converges fast to the unbiased estimates (using our knowledge of the species pool). LONG TIME-SERIES ARE LESS BIASED.


### Loading the data

```{r}

library(tidyverse)

metadata <- read.csv("../../data/simulations/mobsim/IFYAH3130E_jitter_metadata.csv")
load("future_good_solutions.RData")

tst_w_metadata <- inner_join(metadata, tst)

# tst_w_metadata <- tst_w_metadata %>% mutate(facet = paste0("SP: ", S_POOL, " N: ", N_SIM,
#                                          " SAD: ", SAD_COEF, " SIGMA : ", SIGMA))

tst_w_metadata <- tst_w_metadata %>% mutate(facet = paste0(S_POOL, "-", N_SIM,
                                                           "-", SAD_COEF, "-", SIGMA))
```

 \newpage
 
## $T_c$ plots

Remember that $T_c$ measures how fast communities change. We can see here that the communities are changing quite fast (most times at a pace faster that $T_c$). 


```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = raw_T_c)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Characteristic time - raw") +
   geom_smooth()
```
We see above that $T_c$ is a bit biased for small time-series, but asymptotically unbiased. This is so because I used the observed species pool in the timeseries. Once a big proportion of the species in the pool have been recorded, we lost that bias.

```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = T_c)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Characteristic time - complete species pool") +
   geom_smooth()
```
However, if we use our knowledge of the species pool, we get unbiased estimates of $T_c$. 

\newpage

## Asymptotic Jaccard

Asymptotic Jaccard corresponds to the expected Jaccard dissimilarity between two samples far in time, that is, effectively independent. Here we see three things: 

1. A bias in the raw measures.
2. The combination of 200 species in the pool and only 2000 individuals produces very high turnover.
3. The combination of 20 species and 20000 produces small turnover.

```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = raw_a_j)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Asymptotic Jaccard - raw") +
   geom_smooth()
```


```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = a_j)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Asymptotic Jaccard - complete species pool") +
   geom_smooth()
```

\newpage

## Other measures

### Halving time
Akin to Helmut's half-saturation. It is too fast.

```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = raw_h_t)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Halving time - raw") +
   geom_smooth()
```

\newpage

### Affinity
Akin to Helmut's approach. It is less biased than other measures. Low values in the 20 species, 20000 individuals scenario might indicate that there is a fast and small change in these simulations.
```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = raw_Affinity)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Affinity - raw") +
   geom_smooth()
```

\newpage

### Jaccard at the Characteristic time.
Indicates dissimilarity at short time-scales. It is asymptotically unbiased.

```{r, fig.height=6}
ggplot(tst_w_metadata, aes(x = Duration, y = raw_J_t_c)) + geom_point() +
   facet_wrap(~facet, scales = "free_y", ncol = 4) +
   ggtitle("Jaccard at the Characteristic Time - raw") +
   geom_smooth()
```


<!-- Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*. -->

<!-- When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file). -->

<!-- The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed. -->
